Model Finding for (Co)recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-05 tex/main.tex
started making constraints type correct

@@

@@

\newcommand{\boolop}{\oplus}
\newcommand{\forallf}[1]{\forall_{\!#1\:}}
\newcommand{\fnull}{\emptyset}
\newcommand{\vecfarg}[1]{\vec{\concret}_{\vthinspace#1}}
\newcommand{\farg}[1]{\concret_{\vthinspace#1}}
\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{\vthinspace#1}}
\newcommand{\fargsort}[1]{\upsilon_{\vthinspace#1}}

@@

\begin{figure}[t]
\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv( \varphi )$ : 
 \begin{itemize}
   \item[] $\mathsf{let}$ $( \psi, D ) = \conv( \varphi, \ptrue )$;
   %\item[] $\mathsf{assert}( D = \emptyset )$;
   \item[] $\mathsf{return}$ $\psi$
  \end{itemize}
\end{enumerate}
\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv_0( t\typ{\tau}, p )$ : 
 \begin{itemize}
   \item[] $\mathsf{if}$ $\tau \equiv \Bool$ and $t \equiv t_1 \boolop \ldots \boolop t_n$ :
    \begin{itemize}
      \item[] $\mathsf{let}$ $( r_i, D_i ) = \conv_0( t_i, \pol( \boolop, i, p ) )$ for $i = 1, \ldots, n$;
      \item[] $\mathsf{if}$ $p = \ptrue$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge D_1 \wedge \ldots \wedge D_n, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg D_1 \vee \ldots \vee \neg D_n, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n, D_1 \cup \ldots \cup D_n )$
      \end{itemize}
    \end{itemize}
  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
    \begin{itemize}
      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv_0( t_1, p )$;
      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
      \begin{itemize}
        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ), \emptyset )$
      \end{itemize}
      \item[] else
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1, \forall \vec x.\; D_1 )$
      \end{itemize}
    \end{itemize}
   \item[] $\mathsf{else}$
   \begin{itemize}
     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ), f \in \sfundefs{\Sigma} \} )$
   \end{itemize}
 \end{itemize}
\end{enumerate}
\vspace{-2ex}
\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is in definitional form with respect to a set of admissible function definitions $\sfundefs{\Sigma}$.
In the subprocedure $\conv_0$, the procedure $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
where $\boolop$ is an interpreted predicate.
}
\label{fig:encoding}
\end{figure}

@@

The procedure $\conv$ in Figure~\ref{fig:encoding} translates the $\Sigma$-formula
$\varphi$ into an $\extendsig{\Sigma}$-formula.
This calls the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
$\pnone$. It returns a pair $( t'\negvthinspace,\, D )$, where $t'$ is a term of
type $\tau$ and $D$ is a set of formulas encoding additional constraints.
%The role of $D$ is to ensure that
%the (restricted) function definition for $\con{f}$ includes certain tuples in
%its domain, which we explain more in the following.

@@

If $t$ is an application of an interpreted predicate $\boolop
\in \{{\neg}, {\land}, {\lor}, {\teq}, {\lite}\}$,
we recursively call $\conv$ on the arguments $t_i$ and polarity $\pol( \boolop,
i, p )$, where $\pol$ is defined as
\begin{equation*}
\pol( \boolop, i, p ) =
\begin{cases}
\pnone & \text{if $\boolop = {\teq}$ or $(\boolop, i) = (\lite, 0)$} \\
\neg p & \text{if $\boolop = {\neg}$} \\
p & \text{otherwise}
\end{cases}
\end{equation*}
The negation $\neg p$ of polarity $p$ is $\pfalse$ if $p$ is $\ptrue$, $\ptrue$
if $p$ is $\pfalse$, and $\pnone$ if $p$ is $\pnone$.

If $t$ has positive polarity, we construct a conjunction of the result of the
recursive calls with formulas $D_1, \ldots, D_n$. Similarly, if $t$ has
negative polarity, we construct a disjunction of the result of the recursive
calls with the negation of formulas $D_1, \ldots, D_n$. It $t$ has no polarity,
then we return the union of $D_1 \cup \ldots \cup D_n$.

>>>>

%\newcommand\pneg{\neg}
\newcommand\pneg{-}

@@

\def\Chi{\mathrm{X}}

@@

\newcommand{\boolop}{\con{b}}
\newcommand{\forallf}[1]{\forall_{\!#1\:}}
\newcommand{\fnull}{\emptyset}
\newcommand{\vecfarg}[1]{\vec{\concret}_{\vthinspace#1}}
\newcommand{\farg}[1]{\concret_{\vthinspace#1}}
\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{\vthinspace#1}}
\newcommand{\fargsort}[1]{\upsilon_{\vthinspace#1}}

@@

@@

The procedure $\conv$ translates the $\Sigma$-formula
$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
on the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
$\pnone$. It returns a pair $( t'\negvthinspace,\, \Chi )$, where $t'$ is a term of
type $\tau$ and $\Chi$ is a set of formulas encoding additional constraints.
These constraints are necessary if the polarity is $\pnone$; otherwise, $\Chi$
is always the empty set.

\jb{Merge para above with para below}

A translation procedure $\conv$ for a $\Sigma$-formula that is in definitional
form with respect to a set of admissible function definitions
$\sfundefs{\Sigma}$. In the auxiliary procedure $\conv_0$, the procedure $\pol(
\boolop, i, p )$ returns the polarity of the $i${th} argument of a
$\boolop$ application having polarity $p$, where $\boolop$ is an interpreted
predicate.

\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv_0( t\typ{\tau},\, p )$ : 
 \begin{itemize}
   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$:
    \begin{itemize}
      \item[] $\mathsf{let}$ $( r_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
      \item[] $\mathsf{let}$ $\Chi = \Chi_1 \mathrel\cup \cdots \mathrel\cup \Chi_n$;
      \item[] $\mathsf{let}$ $\chi = \bigwedge \Chi$;
      \item[] $\mathsf{if}$ $p = \ptrue$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge \chi,\, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg \chi,\, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n,\, \Chi )$
      \end{itemize}
    \end{itemize}
  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
    \begin{itemize}
      \item[] $\mathsf{let}$ $( r_1,\, D_1 ) = \conv_0( t_1,\, p )$;
      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
      \begin{itemize}
        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
      \end{itemize}
      \item[] else
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1,\, \forall \vec x.\; D_1 )$
      \end{itemize}
    \end{itemize}
   \item[] $\mathsf{else}$
   \begin{itemize}
     \item[] $\mathsf{return}$ $( t,\, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ),\, f \in \sfundefs{\Sigma} \} )$
   \end{itemize}
 \end{itemize}
\end{enumerate}
\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv( \varphi )$ : 
 \begin{itemize}
   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$;
   %\item[] $\mathsf{assert}( D = \emptyset )$;
   \item[] $\mathsf{return}$ $\psi$
  \end{itemize}
\end{enumerate}

%The role of $D$ is to ensure that
%the (restricted) function definition for $\con{f}$ includes certain tuples in
%its domain, which we explain more in the following.

@@

If $t$ is an application of an interpreted predicate $\boolop
\in \{{\neg}, {\land}, {\lor}, {\teq}, {\lite}\}$,
we recursively call $\conv$ on the arguments $t_i$ and polarity $\pol( \boolop,
i, p )$, where $\pol$ is defined as
\begin{equation*}
\pol( \boolop, i, p ) =
\begin{cases}
p & \text{if either $\boolop \in \{{\land}, {\lor}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
\pneg p & \text{if $\boolop = {\neg}$} \\[-\jot]
\pnone & \text{otherwise}
\end{cases}
\end{equation*}
The negation $\pneg p$ of a polarity $p$ is $\pfalse$ if $p$ is $\ptrue$, $\ptrue$
if $p$ is $\pfalse$, and $\pnone$ if $p$ is $\pnone$.

If $t$ has positive polarity, we construct a conjunction of the result of the
recursive calls with the formulas in $D = D_1 \mathrel\cup \cdots \mathrel\cup
D_n$. Similarly, if $t$ has negative polarity, we construct a disjunction of
the result of the recursive calls with the negation of formulas $D_1, \ldots,
D_n$. It $t$ has no polarity, then we return the union of $D_1 \cup \ldots \cup
D_n$.
